Stack Safety as a Security Property

### 2021-03-24 cheri/CheriApproaches.tex
Merged Leo's risc-v work with my most recent and organized the directory

\begin{tabular}{l l l}
  {\tt st}      & Data Store       & Stores data, clearing any capability bits \\
  {\tt ld}      & Data Load        & Loads data, clearing any capability bits \\
  {\tt stc}     & Capability Store & Stores a capability, but failstops rather than store
                                     a local one outside the stack \\
  {\tt ldc}     & Capability Load  & Loads a capability from memory \\
%  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
%                                     decrements the offset \\
%  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
\end{tabular}

@@

Here we will give an example of a call to a function at address {\tt 20} that takes an
argument in register {\tt ra}. Let our call start at address {\tt 0}. Suppose that we have a piece
of code at {\tt 50} that executes a ``landing pad'' for a return. Assume that all capabilities referenced
here are local, and the stack capability is an uninitialized capability that can write locally. There
could be nonlocal capabilities as well, for use in the heap, but to match our model we can also just assume
that such capabilities are derived from a permissive {\tt $ddc}.

>>>>

\begin{tabular}{l l l}
  {\tt st}      & Data Store        & Stores data, clearing any capability bits \\
  {\tt ld}      & Data Load         & Loads data, clearing any capability bits \\
  {\tt stc}     & Capability Store  & Stores a capability, but failstops rather than store
                                      a local one outside the stack \\
  {\tt ldc}     & Capability Load   & Loads a capability from memory \\
  {\tt cca}     & Adjust capability & Adds argument to the capability offset \\ 
%  {\tt incbase} & Increment Base   & Increases the base of a capability by one and
%                                     decrements the offset \\
%  {\tt decbnd}  & Decrement Bound  & Decreases the bound of a capability by one \\
\end{tabular}

@@

Here we will give an example of a call to a function at address {\tt 20} that takes an
argument in register {\tt ra}. Let our call start at address {\tt 0}. Suppose that we have a piece
of code at {\tt 50} that executes a ``landing pad'' for a return. Assume that all capabilities referenced
here are local, and the stack capability is an uninitialized capability that can write locally. There
could be nonlocal capabilities as well, for use in the heap, but to match our model we can also just assume
that such capabilities are derived from a permissive {\tt \$ddc}.

### 2021-03-24 tex/props.tex
Merged Leo's risc-v work with my most recent and organized the directory

    A trace is a potentially infinite, nonempty sequence. We define the append operation \(T \cdot T'\)
    where \(T'\) may be a trace or empty (\(\varepsilon\)):
    \[T_1 \cdot T_2 \triangleq
      \begin{cases}
        T_1 & \text{if } T_2 = \varepsilon \\
        t T_2 & \text{if } T_1 = t \\
        t T_1' \cdot T_2 & \text{if } T_1 = t T_1' \\
      \end{cases}\]

    We identify infinite traces with \(\isinf(T)\). We define a ``trace-of'' operator \(\hookrightarrow\)
    that coinductively relates a machine state with the trace of machine states paired with observations
    produced by repeated application of step:

@@

    Since the step function is total, if \(\mach \hookrightarrow \machT\), then \(\machT\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up through 
    the first state where some condition holds. Let \(f\) be a predicate on machine states.
    Then we define \(\mach \hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
    \(\mach\) until \(f\)''):
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
              \hookrightarrow \machT | f}
           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
    \[\frac{f\ m}
           {m \hookrightarrow (m,\tau)}\]

@@

    \[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
              \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]

    \[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm}
            \neg f\ \mach_1 \hspace{.5cm} \mpstate{\mach_1}{\pol_1} \hookrightarrow_P \machT | f }
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT | f}\]
    \[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm} f\ \mach_1}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs}
            \mpostate{\mach_1}{\pol_1}{\tau} | f}\]

    \COQ{this operation steps until it reaches a state on which \(f\) holds,
    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}

    We project the machine states from either machine trace with \(\pi_m\) and the observations with \(\pi_o\).
    An observational prefix \(\obsT_1 \lesssim \obsT_2\) means that the trace \(\obsT_1\) of observations is
    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations:
    \leo{Point out this is coinductive}

@@

    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
           {\obsT_1 \lesssim \obsT_2}\]
    \leo{Doesn't this case miss the recursive $\obsT_1 \lesssim \obsT_2$ condition?}

    \COQ{This is the same as the coq version minus a redundant case.}

@@

    \COQ{this should be equivalent to the coq version, and we should adopt it.}
    \leo{We had this, but it was much harder to work with for my proofs. We should still
      prove them equivalent if we want to present it this way (which is probably cleaner for
      presentation purposes)}

@@

    \leo{Oh, I like this version. Should make things cleaner in the top part of my proofs
      while leaving the hard low-level invariants unaffected.}

@@

    A stack safety property is a property of a transition system, a policy, and a callmap
    that is decomposed into an integrity property, a confidentiality property, and a safe
    initialization property. We will present eager and observational variants of stack safety.

@@

    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
    and does not interfere with its observable behavior or with the machine state at return. Its
    weaker variant {\it observable stack safety} lifts the latter restriction and states that
    the callee may not leak these secrets directly or by changing the behavior of its caller.

    Finally, {\it safe initialization} deals with the top level outside of a function call, in
    which uninitiaized data should not interfere with any observable behavior.

    \COQ{safe initialization is implemented differently in each property, since
     it's really just confidentiality. But because there's no return it's functionally the same.}

    Our top level properties are {\it eager stack safety} and {\it observable stack safety}:
    \[\begin{split}
      & \textit{observable stack safety} \triangleq \\
      & \textit{observable stack integrity} \land \\
      & \textit{observable stack confidentiality} \land \\
      & \textit{safe initialization}
    \end{split}\]
    \[\begin{split}
      & \textit{eager stack safety} \triangleq \\
      & \textit{eager stack integrity} \land \\
      & \textit{eager stack confidentiality} \land \\
      & \textit{safe initialization}
    \end{split}\]

@@

    \paragraph{Safe Initialization}

    \leo{Do we really need safe initialization? Couldn't we just use the definitions of
      integrity and confidentiality for the toplevel, pointing out that since it never
      returns only confidentiality is interesting? Or would we miss something?}

      In the initial state of a program, nothing has yet been allocated on the stack, and
      there is no guarantee of the initial value of memory in the stack. So there is no
      need to protect a caller's stack data, but reading uninitialized data is unsafe.

@@

      A system enjoys {\it safe initialization} if for any initial states \(\mach_0\)
      and \(\mach_0'\) such that \(\mach_0 \approx_\contour \mach_0'\):

@@

      Above the stack pointer, memory is high integrity in an extension of safe initialization;
      from the perspective of the callee, that memory is not initialized. Below the old stack
      pointer, memory is high integrity and high confidentiality. Registers and the callee's frame
      are low integrity and low confidentiality. These categories inform our integrity and
      confidentiality properties.

>>>>

    A trace is a potentially infinite, nonempty sequence. We define a few operations on traces generically.
    The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
    separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
    \[\frac{f\ t}
           {\mathit{split}(f,tT) = t, tT}\]
    \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
           {\mathit{split}(f,tT) = tT_1,T_2}\]

    The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
    of the first (if it has one) with the first of the second:
    \[\frac{}
           {T_1t \cdot T_2 = T_1 T_2}\]

    Note that both operations are defined coinductively. We identify infinite traces with \(\isinf(T)\).
    We define a ``trace-of'' operator \(\hookrightarrow\) that coinductively relates a machine state with the
    trace of machine states paired with observations produced by repeated application of step:

@@

    Since the step function is total, if \(\mach \hookrightarrow \machT\), then \(\machT\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up through the first state where
    some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
    \(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
    \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
           {\mach \hookrightarrow \machT_1 | f}\]

    %\[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
    %       {\mach \hookrightarrow f | \machT_2}\]

@@

    \[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
              \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]
    \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
           {\mpstatename \hookrightarrow \MPT_1 | f}\]
    %\[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
    %       {\mpstatename \hookrightarrow f | \MPT_2}\]

    We project the machine states from either machine trace with \(\pi_m\) and the observations with \(\pi_o\).
    An observational prefix \(\obsT_1 \lesssim \obsT_2\) means that the trace \(\obsT_1\) of observations is
    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations (once again, coinductively):

@@

    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
           {\obsT_1 \lesssim \obsT_2}\]

@@

@@

@@

    A stack safety property is a property of a transition system, a policy, and a callmap
    that is decomposed into an integrity property and a confidentiality property.
    We will present eager and observational variants of stack safety.

@@

    \sna{Attempt 2...}
    {\it Observable stack confidentiality} states that data outside a callee's stack frame are secret,
    and may not interfere with the observable behavior of the program. That is, the callee's behavior must
    be identical regardless of the configuration of memory outside its frame, and it must not change the
    program state based on that memory in a way that causes future changes in behavior.
    {\it Eager stack confidentiality} additionally states that changes the callee makes to the program state
    must be identical regardless of the configuration of memory.

    Our top level properties are {\it eager stack safety} and {\it observable stack safety}:
    \[\begin{split}
      & \textit{observable stack safety} \triangleq \\
      & \textit{observable stack integrity} \land \\
      & \textit{observable stack confidentiality} \\
    \end{split}\]
    \[\begin{split}
      & \textit{eager stack safety} \triangleq \\
      & \textit{eager stack integrity} \land \\
      & \textit{eager stack confidentiality} \\
    \end{split}\]

    We will separately introduce confidentiality at the top level, to illustrate it simply before
    moving into calls. Integrity at the top level is vacuous because there is nothing to protect.

@@

    \paragraph{Top-level Confidentiality}

    \leo{Do we really need safe initialization? Couldn't we just use the definitions of
      integrity and confidentiality for the toplevel, pointing out that since it never
      returns only confidentiality is interesting? Or would we miss something?}
    \sna{Changed to make it confidentiality, but still introducing contours after. Maybe
         contours should be before?}

      In the initial state of a program, nothing has yet been allocated on the stack, and
      there is no guarantee of the initial value of memory in the stack. So there is no
      need to protect a caller's stack data, but data above the stack frame
      (all of memory) are out of scope and uninitialized, and therefore unsafe to read.

@@

      A system enjoys confidentiality in this case if for any initial states \(\mach_0\)
      and \(\mach_0'\) such that \(\mach_0 \approx_\contour \mach_0'\):

@@

      Above the stack pointer, memory is high integrity; from the perspective of the callee, that
      memory is not initialized. Below the old stack pointer, memory is high integrity and high
      confidentiality. Registers and the callee's frame are low integrity and low confidentiality.
      These categories inform our integrity and confidentiality properties.
